---
title: "GSS challange More Data"
author: "Mya Orozco"
date: "10/01/24"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
dd3 <- mydata %>% group_by(cappun,race) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In the ggplot the attitude of the white race is obviously more in favor. The black and other attitudes are towards opposition. 

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~cappun + race, data=mydata)
rowPerc(table2)
colPerc(table2)
```

For the white group the ratio of favor to oppose is 65.35. Whites of course favor more in their race. As for blacks the ratio is 43.57. In other their ratio is 55:45. 

## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

````{r}
chisq.test(table2)
chisqtestGC(table2)
````

The p-value is 0.00000000000000022 which is very low. So i'm going to run a fisher.t test to check for more data.

## More data version
```{r}
fisherXY(mydata$race, mydata$cappun)
```

In the fisherXY test all the p values were different and are all very low. The p value for white and other is 0.002607. The p value for white and black is 0.00000000000000022. The p value for back and other 0.001564.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Using 3000 variables the Master branch showed us that the white group favors capitol punishment and the black group and the other group oppose capitol punishment. The graphs we made differ in the master branch as we see when we focused our variable to 821 in the number 2 branch so our data had a different conclusion. As the number 2 branch would have a 2:1 ratio of favor for the white group compared to the master branch would have a much closer graph with around 2:3 ratio in favor. The biggest difference is the change with the other group, as they would be seen to have a higher number that oppose capital punishment in the master branch while the number 2 branch they have a higher number in favor of capital difference.